Local Spectral Clustering of Density Upper Level Sets

### 2019-06-17 neurips/local_spectral_density_clustering_arxiv.tex
re-org of Theorem 3

\paragraph{Approximate PPR vector.}  In practice, exactly solving \eqref{eqn:
  ppr_vector} may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\epsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\epsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\epsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \epsilon \Dbf_{uu}\leq
\pbf^{(\epsilon)}(u) \leq \pbf(u).  
\end{equation}
Application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:
  misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\epsilon)}$}. We
formally state and prove this in the appendix.

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
  Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
  \ref{asmp: bounded_volume} for some $r, \sigma, \lambda_{\sigma}, c_0, \gamma
  > 0$ (no bound on maximum density is needed). Then for any $0 < \delta < 1$,
  $\epsilon > 0$, if 
  \begin{equation}
    \label{eqn: conductance_sample_complexity}
    n \geq \frac{(2+\epsilon)^2\log(3/\delta)}{\epsilon^2}\left(\frac{25}
      {6 \lambda_{\sigma}^2\nu(\Csig) \nu_d r^d}\right)^2,
  \end{equation}
  then
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon, 
  \end{equation}
  with probability at least $1-\delta$, where $c_1 > 0$ is a universal constant. 
\end{theorem}

@@

%In this supplement, we present proofs and experimental details for the article ``Local Spectral Clustering of Density Upper Level Sets''.
\section{Proof of Theorem~\ref{}}

\section{Proofs}

Subsections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. In subsections \ref{sec: mixing_time_on_graphs} and \ref{sec: proof_of_proposition_tv_mixing_time}, we establish some results regarding mixing time over general graphs. The proof of Theorem \ref{thm: mixing_time_upper_bound} is then provided in subsections \ref{sec: population_conductance_function} - \ref{sec: proof_of_theorem_mixing_time}. Section~\ref{sec: concentration} gives some general concentration results we use throughout, before we finish with the proofs of Theorems \ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}, and the statement and proof of our results regarding the a\pprspace vector (Corollary \ref{cor: appr}), in subsections \ref{sec: proof_of_misclassification_rate} - \ref{sec: linalg}.

\subsection{Volume estimates}
\label{sec: volume_estimates}

We begin with some notation. Let $\Aset \subseteq \Reals^d$, and for $\sigma \geq 0$, write $\sigma B := B(0,\sigma) = \set{x \in \Rd: \norm{x} \leq \sigma}$ for the closed ball of radius $\sigma$ centered at the origin (and let $B^{\circ}(0,\sigma)$ denote the corresponding open ball). Let $\Asig = \Aset + \sigma B$ be the direct sum of $\Aset$ and $\sigma B$, $\Asig = \set{z = x + y: x \in \Aset, y \in \sigma B}$. Recall that we use $\nu$ for Lebesgue measure, and $\nu_d = \nu(B)$ for $B = (0,1)$.

@@

	Fix $\epsilon > 0$. Our first goal is to find a finite collection $x_1, \ldots, x_N \in \Rd$ (where $N$ is a finite number that may implicitly depend on $\epsilon$) such that
	\begin{equation*}
	\bigcup_{i = 1}^{N} B(x_i, \sigma) \subseteq \Asig \subset \bigcup_{i = 1}^{N} B(x_i, \sigma + \epsilon).
	\end{equation*}
	Note that $\Asig$ is the direct sum of two compact sets, and is therefore itself compact. Moreover, for any $\epsilon > 0$,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists a finite subcover $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation}
	\label{eqn: finite_subcover}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation}
	As a direct consequence of \eqref{eqn: finite_subcover}, $\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon + \delta)$, and by definition for every $x_i \in \Aset$, $B(x_i,\sigma) \in \Asig$. Summarizing our findings, we have
	\begin{equation}
	\label{eqn: finite_subcover-1}
	\bigcup_{i = 1}^{N} B(x_i,\sigma) \subseteq \Asig  ,~\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \delta + \epsilon)
	\end{equation}

>>>>

\sbcomment{You are using $\epsilon$ here and in the future to mean different things. I have
changed this $\epsilon$ to $\varepsilon$.} 
\paragraph{Approximate PPR vector.}  In practice, exactly solving \eqref{eqn:
  ppr_vector} may be too computationally expensive. To address this limitation,
\citet{andersen2006} introduced the \emph{$\varepsilon$-approximate} PPR vector
(aPPR), which we will denote by \smash{$\pbf^{(\varepsilon)}$}. We refer the
curious reader to \citet{andersen2006} for a formal algorithmic definition of
the aPPR vector, and limit ourselves to highlighting a few salient points: the
aPPR vector can be computed in order $\mathcal{O}(1/(\varepsilon\alpha))$ time,
while satisfying the following uniform error bound: 
\begin{equation}
\label{eqn: appr_error}
\textrm{for all $u \in V$}, \quad \pbf(u) - \varepsilon \Dbf_{uu}\leq
\pbf^{(\varepsilon)}(u) \leq \pbf(u).  
\end{equation}
Application of \eqref{eqn: appr_error} within the proofs of Theorems \ref{thm:
  misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}
leads to analogous results which hold for \smash{$\pbf^{(\varepsilon)}$}. We
formally state and prove this in Corollary~\ref{cor: appr} in 
the appendix.

@@

\begin{theorem}
  \label{thm: conductance_upper_bound}
  Fix $\lambda > 0$, and assume $\Cset \in \Cbb_f(\lambda)$ satisfies
  Assumptions \ref{asmp: bounded_density}--\ref{asmp: low_noise_density}, 
  \ref{asmp: bounded_volume} for some $r, \sigma, \lambda_{\sigma}, c_0, \gamma
  > 0$ (no bound on maximum density is needed). Then for any $0 < \delta < 1$,
  $\epsilon > 0$, if 
  \begin{equation}
    \label{eqn: conductance_sample_complexity}
    n \geq \frac{(2+\epsilon)^2\log(3/\delta)}{\epsilon^2}\left(\frac{25}
      {6 \lambda_{\sigma}^2\nu(\Csig) \nu_d r^d}\right)^2,
  \end{equation}
  then with probability at least $1-\delta$, for a universal constant $c_1 > 0$,
  \begin{equation}
    \label{eqn: conductance_additive_error_bound}
    \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon.
  \end{equation}
\end{theorem}

@@

%In this supplement, we present proofs and experimental details for the article ``Local Spectral Clustering of Density Upper Level Sets''.
\section{Proof of Theorem \ref{thm: conductance_upper_bound}}
\label{sec: proof_of_theorem_1}

With this notation in place, the goal of Theorem~\ref{thm: conductance_upper_bound} is roughly to show that, with probability at least $1 - \delta$, for a universal constant $c_1 > 0$ and for any $\epsilon > 0$, for a sufficiently large sample size $n$:
\begin{align*}
\Phi_{n,r}(\Csig[\Xbf]) := \frac{\cut_{n,r}}{\min\{\vol_{n,r}, \vol_{n,r}^c\}} \leq c_1 \frac{d}{\sigma}
    \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
      c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon.
\end{align*}

The proof of this theorem follows essentially from two technical Lemmas. 
The following Lemma relates the terms in the numerator and denominator of $\Phi_{n,r}(\Csig[\Xbf])$ to their expected values.
\begin{lemma}
	\label{lem: prob_bound_cutvol}
	For any $\delta \in (0,1]$,
	\begin{equation}
	\label{eqn: numerator_additive_bound}
	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}}, \\ 
	\frac{\vol_{n,r}}{{n \choose 2}} \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}, \\
	\frac{\vol_{n,r}^c}{{n \choose 2}} \geq p_V^c - \sqrt{\frac{\log(1/\delta)}{n}},
	\end{equation}
	each with probability at least $1 - \delta$. 
\end{lemma}
We note that as a consequence of~\ref{asmp: bounded_volume} we have that $p_V^c \geq p_V$, so it will suffice to lower bound $p_V$ (since a lower bound for $p_V^c$ follows).
The 
following result provides upper and lower bounds on the expected values $p_K$ and $p_V$ respectively:
\begin{lemma}
	\label{lem: expected_density_cut}\label{lem: expected_density_volume}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{align}
	p_K &\leq \frac{4 d \nu_d r^{d+1} \lambda}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig), \label{eqn:claim_one} \\
	p_V &\geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).\label{eqn:claim_two}
	\end{align}
\end{lemma}
%\begin{lemma}
%	\label{lem: expected_density_volume}
%	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
%	\begin{equation*}
%	p_V \geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig).
%	\end{equation*}
%\end{lemma}

Taking these two technical lemmas as given we can now complete the proof of the theorem. 
We fix $\delta \in (0,1]$ and let $\delta' = \delta/3$. We rewrite
\begin{equation}
\label{eqn: conductance_representation_1}
\Phi_{n,r}(\Csig[\Xbf]) = \frac{p_K + \left(\frac{\cut_{n,r}}{{n \choose 2}} - p_K\right)}{p_V + \left(\frac{\min\set{\vol_{n,r}, \vol_{n,r}^c}}{{n \choose 2}} - p_V\right)}.
\end{equation}
By a union bound we see that all three bounds in \eqref{eqn: numerator_additive_bound} hold with probability at least $1 - \delta$. Along with (\ref{eqn: conductance_representation_1}) this yields the upper bound \sbcomment{Changed $\mathbf{X}$ to $X$}
\begin{equation*}
\Phi_{n,r}(\Csig[X]) \leq \frac{p_K + \Err_n}{p_V - \Err_n}
\end{equation*}
for $\Err_n = \sqrt{\frac{\log(1/\delta')}{n}}$.
Now, some straightforward algebraic manipulations yield,
\begin{align*}
\frac{p_K + \Err_n}{p_V - \Err_n} & = \frac{p_K}{p_V} \left(\frac{p_V}{p_V - \Err_n}\right) + \frac{\Err_n}{p_V - \Err_n} \\
& = \frac{p_K}{p_V} + \left(\frac{p_K}{p_V} + 1\right)\frac{\Err_n}{p_V - \Err_n} \\
& \leq \frac{p_K}{p_V} + 2 \frac{\Err_n}{p_V - \Err_n}.
\end{align*}

%The proof of Theorem \ref{thm: conductance_upper_bound} is more or less given by Lemmas \ref{lem: expected_density_cut}, \ref{lem: expected_density_volume}, and \ref{lem: prob_bound_cutvol}. All that remains is some algebra, which we take care of below.

\noindent Using Lemma \ref{lem: expected_density_cut}, we have
\begin{equation*}
\frac{p_K}{p_V} \leq \frac{100rd}{12\sigma} \frac{\lambda}{\lambda_{\sigma}} \frac{\left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right)}{\lambda_{\sigma}}.
\end{equation*}
Then, by the choice of sample size in \eqref{eqn: conductance_sample_complexity}, 
\begin{equation*}
n \geq \frac{(2 + \epsilon)^2 \log\left(\frac{3}{\delta}\right)}{\epsilon^2 p_V^2}
\end{equation*}
which implies $2 \frac{\Err_n}{p_V - \Err_n} \leq \epsilon$, and this yields the desired bound on $\Phi_{n,r}(\Csig[X])$. In order to complete the proof of
the theorem it remains to prove Lemmas~\ref{lem: prob_bound_cutvol} and~\ref{lem: expected_density_cut}.

\subsection{Proof of Lemma~\ref{lem: prob_bound_cutvol}}
We first note that we can write $\cut_{n,r}$ as a double sum,
	\begin{equation}
	\label{eqn: density_cut_expansion}
	\cut_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \not\in \Csig) \1(x_j \in \Csig) \1(\norm{x_i - x_j} \leq r),
	\end{equation}
	and similarly we can write $\vol_{n,r}$ as a sum of indicator functions,
	\begin{equation}
	\label{eqn: volume_expansion}
	\vol_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \in \Csig) \1(x_j \in B(x_i, r)).
	\end{equation}
From this we see that $\cut_{n,r}$ and $\vol_{n,r}$, properly scaled, can be expressed as order-$2$ $U$-statistics,
	\begin{equation*}
	\frac{\cut_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_K(x_i, x_j),~~ \frac{\vol_{n,r}}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_V(x_i, x_j)
	\end{equation*}
	with kernels
	\begin{align*}
	\phi_K(x_i,x_j) & = \1(x_i \in \Csig, x_j \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, x_i \not\in \Csig, \norm{x_i - x_j} \leq r) \\
	\phi_V(x_i,x_j) & = \1(x_i \in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \in \Csig, \norm{x_i - x_j} \leq r). 
	\end{align*}
	Similarly,
	\begin{equation*}
	\frac{\vol_{n,r}^c}{{n \choose 2}} = \frac{1}{{n \choose 2}} \sum_{1 \leq i < j \leq n} \phi_{V^c}(x_i, x_j)
	\end{equation*}
	with kernel,
	\begin{equation*}
	\phi_{V^c}(x_i,x_j) = \1(x_i \not\in \Csig, \norm{x_i - x_j} \leq r) + \1(x_j \not\in \Csig, \norm{x_i - x_j} \leq r). 
	\end{equation*}
	Appealing to a standard concentration inequality for U-statistics (see Lemma \ref{lem: bounded_difference}) we therefore have
	\begin{equation*}
	\frac{\cut_{n,r}}{{n \choose 2}} \leq p_K + \sqrt{\frac{\log(1/\delta)}{n}},~  \frac{\vol_{n,r}}{{n \choose 2}} \geq p_V - \sqrt{\frac{\log(1/\delta)}{n}}, ~ \frac{\vol_{n,r}^c}{{n \choose 2}} \geq p_V^c - \sqrt{\frac{\log(1/\delta)}{n}}
	\end{equation*}
	each with probability at least $1 - \delta$. 
	%The claim follows in light of \ref{asmp: bounded_volume}, which implies $p_V^c \geq p_V$.

\subsection{Proof of Lemma~\ref{lem: expected_density_cut}}
We write $\Pbb(\Aset) = \int_{\Aset} f(x) dx$ for measurable $\Aset \subseteq \Rd$.
We 
let $\Csigr := \set{x: 0 < \dist(x, \Csig) < r}$, where $\Csig$ is as in Theorem \ref{thm: conductance_upper_bound}. 
Our goal will be to upper bound $p_K$ by a term that depends on the probability mass $\Pbb(\Csigr)$, 
and the bulk of our technical effort will be devoted to showing the following upper bound on $\Pbb(\Csigr)$:
%Lemma \ref{lem: expected_number_boundary_points} involves the bulk of the technical effort required to prove Theorem \ref{thm: conductance_upper_bound}; it will be necessary to bound the expected cut size of $\Csig[\Xbf]$ in $G_{n,r}$. 
\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq \frac{2dr}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig)
	\end{equation*}	
\end{lemma}
\noindent Define the \emph{uniform local conductance} $\ell_{\nu,r}(u)$ to be
\begin{equation*}
\ell_{\nu,r}(u) = \nu\bigl(\Csig \cap B(u,r)\bigr).
\end{equation*}
In order to lower bound $p_V$ we will show the following lower bound on the uniform local conductance:
\begin{lemma}
	\label{lem: local_conductance}
	Let $u \in \Csig$. Then, for any $0 < r \leq \frac{\sigma}{2\sqrt{d}}$,
	\begin{equation*}
	\ell_{\nu,r}(u) \geq \frac{6}{25} \nu_d r^d.
	\end{equation*}
\end{lemma}
Taking these two results as given we can now prove each of the two claims~\eqref{eqn:claim_one} and~\eqref{eqn:claim_two} in turn.
\paragraph{Proof of Claim~\eqref{eqn:claim_one}: } Recalling,~\eqref{eqn: density_cut_expansion} 
	and by the linearity of expectations, we obtain
	\begin{equation*}
	p_K = \frac{\mu_K}{{n \choose 2}} = 2 \cdot \Pbb(x_i \not\in \Csig, x_j \in \Csig, \norm{x_i - x_j} \leq r). \tag{for each $i,j$, $i \neq j$}
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_K & = 2 \int_{\Rd \setminus \Csig} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) \dx \\
	& = 2 \int_{\Csigr} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) \dx \\
	& \leq 2 \nu_d r^d \lambda  \int_{\Csigr} f(x) \dx = 2 \nu_d r^d \lambda \Pbb(\Csigr).
	\end{align*}
	where the inequality follows from \ref{asmp: cluster_separation}, which implies $f(x) \leq \lambda$ for $x \in \Csig \setminus \Cset$. Then, upper bounding the integral using \sbcomment{I changed the Lemma referenced here. I think it was wrong before} Lemma \ref{lem: expected_number_boundary_points} gives the final result.

%\subsection{Proof of Lemma~\ref{Z}}
\paragraph{Proof of Claim~\eqref{eqn:claim_two}: } Recalling,~\eqref{eqn: volume_expansion}
%The proof will proceed similarly to Lemma \ref{lem: expected_density_cut}. 
	and by the linearity of expectation we obtain
	\begin{equation*}
	p_V = \frac{\mu_V}{{n \choose 2}} = 2 \cdot \Pbb(x_i \in \Csig, x_j \in B(x_i,r)). \tag{for any $i,j$, $i \neq j$. }
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_V & = 2 \int_{\Csig} f(x) \Pbb(B(x,r)) \dx \\
	& \geq 2 \int_{\Csig} f(x) \Pbb(B(x,r) \cap \Csig) \dx
	\end{align*}
	whence the claim then follows by Lemma \ref{lem: local_conductance}.
It remains to prove Lemmas~\ref{lem: expected_number_boundary_points} and~\ref{lem: local_conductance}, and we turn our attention to this next.
\subsection{Proof of Lemma~\ref{lem: expected_number_boundary_points}}
The proof of this Lemma relies on certain volume estimates whose statement and proof we defer to Appendix~\ref{sec: volume_estimates}.
	We partition $\Csigr$ into slices based on distance from $\Csig$ as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Csigr: t_{i,k} < \frac{\dist(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k},
	\end{equation*}
	\sbcomment{Changed $t_i$ to $t_{i,k}$.}
	where $t_{i,k} = i/k$ for $i = 0, \ldots, k - 1$. As a result, for any $k \in \mathbb{N}$,
	\begin{equation}
	\label{eqn: partition_ub}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) \dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) \dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation}
	Assumptions~\ref{asmp: bounded_density} and \ref{asmp: low_noise_density} imply the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma},
	\end{equation*}
	and writing
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) =: \nu_{i+1,k} - \nu_{i,k},
	\end{equation*}
	we have
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \underbrace{\sum_{i = 1}^{k} 
	\nu_{i,k} \biggl( \left[\lambda_{\sigma} - c_0(rt_{i-1,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - c_0(rt_{i,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - c_0r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi}
	\end{align}
	\sbcomment{I changed the braces above -- I think they were incorrect earlier} where the second equality comes from rearranging terms in the sum.

	% AJG 5/14/19: This argument makes sense, correct?
	We first consider the term $\Sigma_k$. $\Cset$ has finite diameter by Assumption~\ref{asmp: bounded_density}, as otherwise $\int_{\Csig} f(x) dx = \infty$. Letting $\overline{\Cset}$ be the closure of $\Cset$, we observe that $\overline{\Csig} = \overline{\Cset} + \sigma B$, and moreover for any $\delta > 0$, $\nu(\overline{\Csig} + \delta B) = \nu(\Csig + \delta B)$ (as the boundary $\partial(\Csig + \delta B)$ is Lipschitz and therefore has measure zero). As a result, for each $t_{i,k}, i = 1, \ldots,k$ we may apply Lemma \ref{lem: expansion_volume} to $\overline{\Cset}$ and obtain
	\begin{equation}
	\label{eqn: slice_volume_bound}
	\nu_{i,k} = \nu(\Csig + rt_{i,k}B) \leq \nu(\Csig)\left(1 + \frac{rt_{i,k}}{\sigma}\right)^d
	\end{equation}
	which in turn gives
	\begin{align}
	\Sigma_k & \leq c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ rt_{i,k}}{\sigma}\right)^d \biggl( (t_{i,k})^{\gamma} - (t_{i-1,k})^{\gamma}\biggr) \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}).~~~~~~~~~~~~~~ (\text{substituting}~u_{i,k} := t_{i,k}^{\gamma}) \label{eqn: Sigmak_riemann_sum}
	\end{align}
	The expression in~\eqref{eqn: Sigmak_riemann_sum} is a Riemann sum, and taking the limit as $k \to \infty$ we obtain
	\begin{align}
	\lim_{k \to \infty} c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}) & = c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{r u^{1/\gamma}}{\sigma}\right)^{d} du \nonumber \\
	& \overset{\text{(i)}}{\leq} c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{2 d r u^{1/\gamma}}{\sigma}\right) du \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \left(1 + \gamma \frac{2 d r}{(\gamma + 1)\sigma}\right). \label{eqn: Sigmak_integral}
	\end{align}
	where $\text{(i)}$ follows from the upper bound in Lemma \ref{lem: Taylor_series} in light of the fact $r \leq \sigma / 2d$.

	An upper bound on $\xi$ follows from largely the same logic, although it does not involve integration:
	\begin{align}
	\xi & \overset{\text{(ii)}}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{ r}{\sigma}\right)^d(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} \nonumber \\
	& \overset{\text{(iii)}}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{2d r}{\sigma}\right)(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} = \nu(\Csig) \biggl\{ \frac{2dr}{\sigma}(\lambda_{\sigma} - c_0r^{\gamma}) - c_0 r^{\gamma} \biggr\}. \label{eqn: xi_ub}
	\end{align}
	where $\text{(ii)}$ follows from \eqref{eqn: slice_volume_bound}, and $\text{(iii)}$ from Lemma \ref{lem: Taylor_series}. As the bounds in \eqref{eqn: partition_ub} and \eqref{eqn: telescoping_sum} hold for all $k$, these along with \eqref{eqn: Sigmak_integral} and \eqref{eqn: xi_ub} imply the desired result.

%\subsection{Proof of Lemma~\ref{lem: expected_number_boundary_points}}

\subsection{Proof of Lemma~\ref{lem: local_conductance}}
The proof of this result relies on estimates of the volume of spherical caps which we defer to Appendix~\ref{sec:caps}.
	Since $u \in \Csig$ there exists $x \in \Cset$ such that $u \in B(x, \sigma)$, and as $B(x,\sigma) \subseteq \Csig$,
	\begin{equation*}
	\nu\bigl(B(u, r) \cap B(x, \sigma)\bigr) \leq \nu\bigl(B(u, r) \cap \Csig \bigr)
	\end{equation*}
	Without loss of generality, let $\norm{u - x} = \sigma$; it is not hard to see that if $\norm{u - x} < \sigma$, the volume of the overlap will only grow. Then, since $\norm{u  - x} = \sigma$, $B(u, r) \cap B(x, \sigma)$ contains a spherical cap of radius $r$ and height
	\begin{equation*}
	h = r - r^2/2\sigma = r \left( 1 - \frac{r}{2 \sigma} \right)
	\end{equation*}	
	which by Lemma \ref{lem: volume_of_spherical_cap} has volume
	\begin{equation*}
	\nu_{\text{cap}} = \frac{1}{2} \nu_d r^d I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right)
	\end{equation*}
	with $\alpha = 1 - \frac{2rh - h^2}{r^2} = \frac{r^2}{4 \sigma^2} \leq \frac{1}{16d}$. 
	By Lemmas \ref{lem: beta_integral} (applied with $t = 1$) and \ref{lem: beta_function},
	\begin{align*}
	I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right) & \geq 1 - \frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \frac{3}{4\sqrt{d}} \\
	& \geq 1 - \frac{3}{4}\sqrt{\frac{d+2}{\pi d}} \geq 1 - \frac{3}{4}\sqrt{\frac{3}{2 \pi}}.
	\end{align*}

\subsection{Volume Estimates}
\label{sec: volume_estimates}
%\section{Proofs}
%
%Subsections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. In subsections \ref{sec: mixing_time_on_graphs} and \ref{sec: proof_of_proposition_tv_mixing_time}, we establish some results regarding mixing time over general graphs. The proof of Theorem \ref{thm: mixing_time_upper_bound} is then provided in subsections \ref{sec: population_conductance_function} - \ref{sec: proof_of_theorem_mixing_time}. Section~\ref{sec: concentration} gives some general concentration results we use throughout, before we finish with the proofs of Theorems \ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}, and the statement and proof of our results regarding the a\pprspace vector (Corollary \ref{cor: appr}), in subsections \ref{sec: proof_of_misclassification_rate} - \ref{sec: linalg}.

%\subsection{Volume estimates}

We begin by recalling some notation. We let $\Aset \subseteq \Reals^d$, and for $\sigma \geq 0$, write $\sigma B := B(0,\sigma) = \set{x \in \Rd: \norm{x} \leq \sigma}$ for the closed ball of radius $\sigma$ centered at the origin (and let $B^{\circ}(0,\sigma)$ denote the corresponding open ball). Let $\Asig = \Aset + \sigma B$ be the direct sum of $\Aset$ and $\sigma B$, $\Asig = \set{z = x + y: x \in \Aset, y \in \sigma B}$. Recall that we use $\nu$ for Lebesgue measure, and $\nu_d = \nu(B)$ for $B = (0,1)$.

@@

	Fix $\epsilon > 0$. Our first goal is to find a finite collection $x_1, \ldots, x_N \in \Rd$ (where $N$ is a finite number that may implicitly depend on $\epsilon$) such that
	\begin{equation*}
	\bigcup_{i = 1}^{N} B(x_i, \sigma) \subseteq \Asig \subset \bigcup_{i = 1}^{N} B(x_i, \sigma + \epsilon).
	\end{equation*}
	Note that $\Asig$ is the direct sum of two compact sets, and is therefore itself compact. Moreover, for any $\epsilon > 0$,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists a finite subcover $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation}
	\label{eqn: finite_subcover}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation}
	As a direct consequence of \eqref{eqn: finite_subcover}, $\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon + \delta)$, and by definition for every $x_i \in \Aset$, $B(x_i,\sigma) \in \Asig$. Summarizing our findings, we have
	\begin{equation}
	\label{eqn: finite_subcover-1}
	\bigcup_{i = 1}^{N} B(x_i,\sigma) \subseteq \Asig  ,~\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \delta + \epsilon).
	\end{equation}
\noindent	We next show a lower bound on $\nu(\Asig)$. Partition $\Asig$ using the balls $B(x_i,\sigma)$, meaning let $\Aset_{\sigma}^{(1)} := B(x_1,\sigma)$, $\Aset_{\sigma}^{(2)} := B(x_2,\sigma) \setminus B(x_1,\sigma)$, and so on, so that
	\begin{equation*}
	\Aset_{\sigma}^{\text{(i)}} := B(x_i,\sigma) \setminus \bigcup_{j = 1}^{i - 1} \Aset_{\sigma}^{(j)}. \tag{$i = 1,\ldots,N$}
	\end{equation*}
	Observe that $\bigcup_{i = 1}^{N} \Asig^{\text{(i)}} = \bigcup_{i = 1}^{N} B(x_i,\sigma)$, so by \eqref{eqn: finite_subcover} $\Asig \supseteq \bigcup_{i = 1}^{N} \Asig^{\text{(i)}}$. As $\Asig^{(1)},\ldots, \Asig^{(N)}$ are non-overlapping,
	\begin{align*}
	\nu(\Asig) & \geq \sum_{i = 1}^{N} \nu(\Asig^{\text{(i)}}) \\
	& = \sigma^d \nu_d \sum_{i = 1}^{N}  \frac{\nu(\Asig^{\text{(i)}})}{\nu(B(x_i,\sigma))}
	\end{align*}
	We turn to proving an analogous upper bound on $\nu(\Asig + \delta B)$. Let $\Aset_{\sigma + \epsilon + \delta}^{(1)} := B(x_1,\sigma + \delta + \epsilon)$ and
